---
title: "03_describe.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

# Part 03: describe

## Import the required packages

```{r}

#| label: load packages
#| echo: false
#| message: false

library("tidyverse")

```
## Describing the data tables

After cleaning and augmenting the data three tables will be present. One table containing the regulators and target molecules from before and after treatment, a second table containing the target regulators and the  ..... 



## Describing of dimensions of the data

The data tables each consists of:

Table X: X observations, and X variables. 
Table XX: X observations, and X variables.
Table XXX: X observations, and X variables.

This can be analyzed with the code below.

### Dimensions of table X 

```{r}

#| label: data dimensions
#| echo: false


# table1_dimensions <- dim()
# table1_variables <- ncol()
# table1_observations <- nrow()

```

### Dimensions of table XX
```{r}

#| label: data dimensions
#| echo: false


# table2_dimensions <- dim()
# table2_variables <- ncol()
# table3_observations <- nrow()

```

### Dimensions of table XXX
```{r}

#| label: data dimensions
#| echo: false


# table3_dimensions <- dim()
# table3_variables <- ncol()
# table4_observations <- nrow()

```
