---
title: "03_describe.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

# Part 03: describe

## Import the required packages

```{r}

#| label: load packages
#| echo: false
#| message: false

library("tidyverse")

```

## Contents the data tables

After cleaning and augmenting the data, three tables will be present: (I) A table containing the regulator molecules and their respective expression levels, both before and after treatment, called XXXXX. (II) A table containing the regulator molecules and their related target molecules, again before and after treatment, called XXXX. (III) Lastly, a table containing the co-expression modules, and the regulator molecules that play a role in each of these modules, called XXXX.

These tables can be found as .tsv files, generated by the previous code.

## Dimensions of the data

The data tables each consists of:

Table X: X observations, and X variables. Table XX: X observations, and X variables. Table XXX: X observations, and X variables.

This can be analyzed with the code below.

### Dimensions of table X

```{r}

#| label: data dimensions
#| echo: false


# table1_dimensions <- dim()
# table1_variables <- ncol()
# table1_observations <- nrow()

```

### Dimensions of table XX

```{r}

#| label: data dimensions
#| echo: false


# table2_dimensions <- dim()
# table2_variables <- ncol()
# table3_observations <- nrow()

```

### Dimensions of table XXX

```{r}

#| label: data dimensions
#| echo: false


# table3_dimensions <- dim()
# table3_variables <- ncol()
# table4_observations <- nrow()

```
