---
title: "05_analysis_1"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 05: analysis 1

The first analysis is that of the genetic expression levels. From the .tsv file xxxxxx, significantly up- and down-regulated genes are depicted below.

#### Install the required packages

```{r}

#| label: load packages
#| message: false
#| echo: false

library("tidyverse")
library("ggrepel")

```

#### Generate a vulcano plot of the gene expression

To depict which genes are up- and down-regulated, a volcano plot is made. The differences in colors indicate the genes present before and after DMARD treatment.

```{r}

#| label: volcano plot
#| echo: false

# Import 03_dat_aug.tsv file |>
  ggplot(aes(x = # expr. log ratio,
             y = # p.value,
             colour = # before vs after treatment,
             label = # up regulated ?? )) +
  geom_point(size = 1,
             alpha = 0.5) +
  #geom_text_repel(size = 4,
                  #max.overlaps = 20) + These two lines might not be necessary
  geom_hline(yintercept = 0) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(hjust = 1),
    plot.subtitle = element_text(hjust = 1),
    legend.position = "none") +
  labs(
    x = "log ratio of gene expression",
    y = "p.value of expression",
    title = "Genetic regulation, related with rheumatoid arthritis (RA)",
    subtitle = "Genes highlighted in turquoise, depict those expressed after treatment",
    caption = "Data from DOI: 10.1371/journal.pone.0183928")

```


