---
title: "Part 01: Load"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

The first part of the project workflow is to load the raw data into the project workflow and import the required packages.

### Import the required packages

```{r}
#| label: load packages
#| message: false

library("tidyverse")
library("readxl")

```

### Load the data

The data, used for this project, is downloaded from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5580991/. More specifically; supplementary table 1 and 3.

Data table 1 contains information on co-expression modules. Each column is a group (called module) of co-expressed genes. Each cell underneath the column header contains the symbol of a gene that is part of this respective module. A total of 20 modules of co-expressed genes is defined. Data table 3 consists of data on predicted regulators. Each row is a regulator molecule, and of each molecule, characteristics such as the expression level and molecule type are described. A total of 849 regulator molecules are described.

When loading the data, blank fields are replaced with "NA".

```{r}
#| label: load data
#| message: false

coexpression_raw <- read_excel("../data/_raw/pone.0183928.s001.xlsx",
                               na = c(" ",
                                      "NA"))

regulators_before_raw <- read_excel("../data/_raw/pone.0183928.s003.xlsx",
                                    na = c(" ",
                                           "NA"))

regulators_after_raw <- read_excel("../data/_raw/pone.0183928.s003.xlsx",
                                   2,
                                   na = c(" ",
                                          "NA"))

```

### Convert the data files to the correct format

In order to enable data analysis, the original .xlsx files are converted to .tsv and saved in the data folder in the project.

```{r}
#| label: save data

write_tsv(coexpression_raw,
           "../data/01_coexpression_load.tsv.gz",
           col_names = TRUE)

write_tsv(regulators_before_raw,
          "../data/01_regulators_before_load.tsv.gz",
          col_names = TRUE)

write_tsv(regulators_after_raw,
          "../data/01_regulators_after_load.tsv.gz",
          col_names = TRUE)
```
