---
title: "02_clean.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 02: clean

### Accessing relevant functions

In order to perform the required cleaning steps on the acquired data frames, an in order to avoid repetition of the code, functions were created. These can be found in the file mentioned below.

```{r}
#| label: use functions from file

source("99_proj_func.R")
```


#### Import the required packages

```{r}
#| label: load packages
#| message: false

library("tidyverse")
library("patchwork")

```

### Read in the loaded data files

```{r}
#| label: load in data

coexpression_clean <- read_tsv("../data/01_coexpression_load.tsv.gz",
                               show_col_types = FALSE)

regulators_before_clean <- read_tsv("../data/01_regulators_before_load.tsv.gz",
                                    show_col_types = FALSE, skip = 1)

regulators_after_clean <- read_tsv("../data/01_regulators_after_load.tsv.gz",
                                   show_col_types = FALSE, skip = 1)
```

### Cleaning of co-expression data

As stated before, data table 1 contains data on 20 co-expression modules. In the code below, the table is tidied, so that each column represents a single module, and the row lists all genes within that given module.


#### Cleaning of table 1

#### Cleaning of table 3

##### Before treatment

The headers of the regulators data are corrected and the column called "Mechanistic Network" is excluded from the data, since it is not relevant for further analysis.



##### After treatment

As with the data regulators from before treatment the headers are corrected and the column called "Mehcanistic Network" is removed.


##### Joining of expression data



##### Joining of target data


##### Making tsv files

