---
title: "03_augment.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 03: augment

The third part of the project is to augment the cleaned data.

### Import required packages

```{r}
#| label: load packages
#| echo: false

library("tidyverse")

```

The cleaned data files are loaded in.

```{r}

#| label: load data

coexpression_aug <- read_tsv("../data/02_coexpression_clean.tsv.gz",
                               show_col_types = FALSE)
expr_aug <- read_tsv("../data/02_expr_clean.tsv.gz",
                     show_col_types = FALSE)
target_aug <- read_tsv("../data/02_target_clean.tsv.gz",
                       show_col_types = FALSE)

```

### Augment the dataframes

No new variables were added to the dataframe containing the coexpression data.

#### Augment the dataframe containing gene expression levels

A variable indicating whether a regulator was up- or down-regulated both before and after treatment was added.

```{r}

#| label: add column showing up- or downregulation 

expr_aug <- expr_aug |>
  mutate(regulation = case_when( expr_log_ratio_before < expr_log_ratio_after ~ "Upregulated",
                                 expr_log_ratio_before > expr_log_ratio_after ~ "Downregulated"))
```

Furthermore, a variable that shows the level of change in gene expression for the regulators was added.

```{r}
#| label: add change in regulation column 

expr_aug <- expr_aug |>
  mutate(change_in_regulation = expr_log_ratio_after - expr_log_ratio_before)

```

#### Augment the dataframe containing the target molecules 

A variable showing the difference in target molecules after treatment compared to before treatment for each regulator was added.

```{r}
target_aug <- target_aug |> 
  mutate(change_in_n_molecules = n_after - n_before) |> 
  relocate(change_in_n_molecules,
           .after = n_after)
```

#### Saving the augmented data frames in new files

Lastly, the augmented dataframes were saved as zipped tsv files.

```{r}
#| label: tsv files

write_tsv(coexpression_aug,
          "../data/03_coexpression_aug.tsv.gz")

write_tsv(expr_aug,
          "../data/03_expr_aug.tsv.gz")

write_tsv(target_aug,
          "../data/03_target_aug.tsv.gz")

```
