---
title: "03_augment.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 03: augment

#### Import libraries 

```{r}

#| label: load packages
#| echo: false

library("tidyverse")

```

#### Load in the clean data files 

```{r}

coexpression_aug <- read_tsv("../data/02_coexpression_clean.tsv",
                               show_col_types = FALSE)
expr_aug <- read_tsv("../data/02_expr_clean.tsv",
                     show_col_types = FALSE)
target_aug <- read_tsv("../data/02_target_clean.tsv",
                       show_col_types = FALSE)

```

#### No changes are made to the dataframe containing the coexpression data

#### Augmenting the dataframe reflecting the gene expression levels

Adding a variable that indicates weather a regulator was up- or down-regulated,
before and after treatment 

```{r}

#| label: add upgerulated/downregulated column 

expr_aug <- expr_aug |>
  mutate(regulation = case_when( expr_log_ratio_before < expr_log_ratio_after ~ "Upregulated",
                                 expr_log_ratio_before > expr_log_ratio_after ~ "Downregulated"))
```

Adding a varaible that reflects the level of change in regulation, the
difference between before and after 
```{r}

#| label: add upgerulated/downregulated column 

expr_clean |> 
  mutate(Regulation = case_when("Expr Log Ratio (before)" < "Expr Log Ratio (after)" ~ "Upregulated",
                                "Expr Log Ratio (before)" > "Expr Log Ratio (after)" ~ "Downregulated"))

```

#### Augmenting the dataframe reflecting the target molecules

```{r}

expr_aug <- expr_aug |>
  mutate(change_in_regulation = expr_log_ratio_after - expr_log_ratio_before)

```

#### Augmenting the dataframe reflecting the target molecules 

Adding the number of target molecules before treatment per regulator

```{r}

target_aug <- target_aug |> 
  rowwise() |> 
  mutate(n_before = sum(!is.na(c_across(contains("before"))))) |> 
  relocate("n_before",
           .after = "upstream_regulator")

```

Adding the number of target molecules after treatment per regulator

```{r}

target_aug <- target_aug |> 
  rowwise() |> 
  mutate(n_after = sum(!is.na(c_across(contains("after"))))) |> 
  relocate("n_after",
           .after = "n_before")

```

Saving the augmented dataframes as tsv files

```{r}

write_tsv(coexpression_aug,
          "../data/03_coexpression_aug.tsv")

write_tsv(expr_aug,
          "../data/03_expr_aug.tsv")

write_tsv(target_aug,
          "../data/03_target_aug.tsv")

```
