---
title: "03_augment.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 03: augment

Import libraries 
```{r}

#| label: load packages
#| echo: false

library("tidyverse")

```

Load in the clean data files 
```{r}

#| label: load data

coexpression_aug <- read_tsv("../data/02_coexpression_clean.tsv.gz",
                               show_col_types = FALSE)
expr_aug <- read_tsv("../data/02_expr_clean.tsv.gz",
                     show_col_types = FALSE)
target_aug <- read_tsv("../data/02_target_clean.tsv.gz",
                       show_col_types = FALSE)

```

#### No changes are made to the dataframe containing the coexpression data

#### Augmenting the dataframe reflecting the gene expression levels

Adding a variable that indicates weather a regulator was up- or down-regulated,
before and after treatment 
```{r}

#| label: add upgerulated/downregulated column 

expr_aug <- expr_aug |>
  mutate(regulation = case_when( expr_log_ratio_before < expr_log_ratio_after ~ "Upregulated",
                                 expr_log_ratio_before > expr_log_ratio_after ~ "Downregulated"))
```

Adding a varaible that reflects the level of change in regulation, the
difference between before and after 
```{r}
#| label: add change in regulation column 

expr_aug <- expr_aug |>
  mutate(change_in_regulation = expr_log_ratio_after - expr_log_ratio_before)

```

```{r}
target_aug <- target_aug |> 
  mutate(change_in_n_molecules = n_after - n_before) |> 
  relocate(change_in_n_molecules,
           .after = n_after)
```

  
#### Saving the augmented data frames as zipped tsv files

```{r}
#| label: tsv files

write_tsv(coexpression_aug,
          "../data/03_coexpression_aug.tsv.gz")

write_tsv(expr_aug,
          "../data/03_expr_aug.tsv.gz")

write_tsv(target_aug,
          "../data/03_target_aug.tsv.gz")

```


