---
title: "03_describe.qmd"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

# Part 04: describe

## Import the required packages

```{r}
#| label: load packages
#| message: false

library("tidyverse")
library("ggthemes")

```

## Description of the data tables

After cleaning and augmenting the data, 3 main dataframes will be present:

-   A table containing coexpression modules and the genes for the regulator molecules that play a role in each of these modules called "03_coexpression_aug"
-   A table containing the regulator molecules and their related target molecules, both before and after treatment called "03_expr_target_aug"
-   A table containing the regulator molecules and their respective expression levels, both before and after called "03_expr_intersect_aug"

Furthermore, the dataframe

These tables can be found as .tsv.gz files, generated by the previous code.

## Dimensions of the data

The data tables each consists of:

Table X: X observations, and X variables. Table XX: X observations, and X variables. Table XXX: X observations, and X variables.

This can be analyzed with the code below.

### Dimensions of table coexpression_aug

```{r}

#| label: data dimensions
#| echo: false


table1_dimensions <- dim(coexpression_aug)
table1_variables <- ncol(coexpression_aug)
table1_observations <- nrow(coexpression_aug)

table1_dimensions
table1_variables
table1_observations

```

The regulator molecules in the data table are categorized by molecule type. The code below generates a box plot, depicting the number of regulator molecules in each of the defined types.

```{r}

#| label: box plot of molecule types
#| echo: false


```

### Dimensions of table II

```{r}

#| label: data dimensions
#| echo: false


table2_dimensions <- dim(expr_)
table2_variables <- ncol()
table3_observations <- nrow()

```

Each regulator molecule regulates a certain number of genes. The code below generates a box plot, showing the distribution of the number of genes per regulator molecule.

```{r}

#| label: box plot of genes per regulator

```

### Dimensions of table III

```{r}

#| label: data dimensions
#| echo: false


# table3_dimensions <- dim()
# table3_variables <- ncol()
# table3_observations <- nrow()

```

### Barplots

```{r}
# barplot for target molecules before treatment

ggplot(
  # Setting data
  data = target_intersect_aug,
  # Sorting it after highest frequency
  aes(x = upstream_regulator, y = n_before)
  )+ 
  # Making a barplot
  geom_col()+
  # Setting titles
  labs(
    title = "Target molecules before treatment",
    subtitle = "The number of times a target molecule is present in the dataset before treatment",
    x = "Target molecule",
    y = "Count"
  )+
  # colorblind friendly colors
  scale_color_colorblind()

```

```{r}
# barplot for target molecules after treatment

ggplot(
  # Setting data
  data = target_intersect_aug,
  # Sorting it after highest frequency
  aes(x = upstream_regulator, y = n_after)
  )+ 
  # Making a barplot
  geom_col()+
  # Setting titles
  labs(
    title = "Target molecules before treatment",
    subtitle = "The number of times a target molecule is present in the dataset before treatment",
    x = "Target molecule",
    y = "Count"
  )+
  # colorblind friendly colors
  scale_color_colorblind()
```
