---
title: "05_analysis_1"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

## Part 05: analysis 1

The first analysis is that of the genetic expression levels. From the .tsv file "03_expr_aug.tsv.gz", significantly up- and down-regulated genes are depicted below.

#### Load the needed .tsv file

```{r}
#| label: load file

gene_expr <- read_tsv("../data/03_expr_aug.tsv.gz",
                      show_col_types = FALSE)

```


#### Load the required packages

```{r}

#| label: load packages
#| message: false
#| echo: false

library("tidyverse")
library("ggrepel")
library("patchwork")

```

#### Generate a vulcano plot of the gene expression

To depict which genes are up- and down-regulated, a volcano plot is made. The differences in colors indicate the genes present before and after DMARD treatment.

```{r}

#| label: volcano plot
#| echo: false

 plot_before <- gene_expr |>
  ggplot(aes(x = expr_log_ratio_before,
             y = p_value_of_overlap_before,
             colour = regulation,
             label = upstream_regulator)) +
  #geom_vline(xintercept = 0) +
  #geom_label() +
  geom_point(size = 1,
             alpha = 0.5) +
  geom_text_repel(size = 3,
                  max.overlaps = 50) +
  geom_hline(yintercept = 0) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "none") +
  labs(
    x = "log ratio of gene expression",
    y = "p.value of expression",
    title = "Genetic expression in RA patients",
    subtitle = "Before DMARD treatment",
    caption = "Data from DOI: 10.1371/journal.pone.0183928")

```

```{r}

#| label: volcano plot
#| echo: false

 plot_after <- gene_expr |>
  ggplot(aes(x = expr_log_ratio_after,
             y = p_value_of_overlap_after,
             colour = regulation, 
             label = upstream_regulator)) +
  #geom_vline(xintercept = 0) +
  #geom_label() +
  geom_point(size = 1,
             alpha = 0.5) +
  geom_text_repel(size = 3,
                  max.overlaps = 50) +
  geom_hline(yintercept = 0) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "none") +
  labs(
    x = "log ratio of gene expression",
    y = "p.value of expression",
    title = "Genetic expression in RA patients",
    subtitle = "After DMARD treatment",
    caption = "Data from DOI: 10.1371/journal.pone.0183928")

```


```{r}
plot_before
```
```{r}
plot_after
```

