---
title: "Part 06: analysis 2"
format: html
editor: visual
author: "Anna Boateng: s175562, Marlies Goedknegt: s220675, Anne Gjelstrup: s194530, Katja Jagd: s185395, Olivia Ronnenberg: s183359"
---

The sixth part of the project is the second analysis, where ...

### Accessing relevant functions

```{r}
#| label: use functions from file

source("99_proj_func.R")
```

### Import required packages

```{r}
#| label: load packages
#| message: false

library("tidyverse")
```

### Load in datafiles

```{r}
#| label: load augmented data files

coexpression_analysis <- read_tsv("../data/03_coexpression_aug.tsv.gz",
                               show_col_types = FALSE)

target_intersect_analysis <- read_tsv("../data/03_target_intersect_aug.tsv.gz",
                       show_col_types = FALSE)
```

### Number of target molecules before and after treatment

The upstream regulators each influence several target molecules. To investigate whether this influence was affected by the treatment, a plot was created to indicate the change of the number of target molecules before and after treatment.

```{r}
#| label: generate target molecules plot

targets_plot <- target_intersect_analysis |> 
  ggplot(aes(y = upstream_regulator,
             x = mol_change,
             fill = change_class)) +
  geom_col(color  = "grey", 
           alpha = 0.5) +
  scale_fill_manual(values = c("Decrease" = "#F8766D",
                             "Increase" = "#00BFC4")) +
  theme_minimal(base_size = 12) +
  theme(axis.text.y=element_text(size=7),
        legend.title=element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(y = "Upstream regulator",
       x = "Change in number of target molecules",
       title = "Change in the number of target molecules",
       subtitle = "Comparing after treatment to before") + 
  scale_x_continuous(breaks = c(-15,-10, -5, 0, 5, 10, 15)) +
  coord_fixed(ratio=4/3) + 
  geom_vline(xintercept = 0,
             alpha = 0.5,
             color = "grey")

targets_plot

ggsave(filename = "06_targets_plot.png",
       plot = last_plot(),
       device = "png",
       path = "../results",
       scale = 1,
       limitsize = TRUE)


```

The plot indicates that the number of target molecules, influenced by each upstream regulator, changes after treatment. There are cases of both increasing, decreasing and unchanged number of target molecules. This highlights the shift in the behavior of the upstream regulators after treatment, which is relevant information when assessing the effect of the treatment, its possible benefits and side effects.

### Biological context of the target molecules

The analysis of the expression data, together with the target molecule analysis above, revealed the following interesting regulators: S100A9, TGFB1, and TLR2. These regulators could be potential drug targets.

As mentioned, the modules in the co-expression data set represent possible biological pathways. Therefore a further analysis was performed to look into which modules the target molecules of the candidates regulators belonged to.

By using the function "module_counter()" on each candidate it was possible to connect the information from the co-expression modules with the target molecules of the regulators. The function counts how many target molecules belongs to each module.

```{r}
#| label: define the selected molecule candidates

candidate_1 <- module_counter(target_intersect_analysis, "S100A9")
candidate_2 <- module_counter(target_intersect_analysis, "TGFB1")
candidate_3 <- module_counter(target_intersect_analysis, "TLR2")
```

The results from all of the candidates was merged into a small data set called "module_analysis".

```{r}
#| label: create data set for the module analysis

module_analysis <- candidate_1 |> 
  full_join(candidate_2,
            by = 'modules') |> 
  full_join(candidate_3,
            by = 'modules') |> 
  mutate_if(is.numeric,
            replace_na,
            0) |> 
  pivot_longer(cols = -modules,
               names_to = "candidate",
               values_to = "counts") |> 
  arrange(as.integer(str_extract(modules, "\\d+")))
```

The result of the module analysis is visualized in the bar plot below.

```{r}
#| label: generate module analysis plot

modules_plot <- module_analysis |>
  ggplot(aes(x = fct_inorder(modules), y = counts, fill = candidate)) +
  geom_bar(stat = "identity",
           color = "grey",
           alpha = 0.5,
           position = position_dodge(preserve = "single"),
           ) +
  geom_hline(yintercept = 0) + 
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 20,
                                   hjust = 1,
                                   vjust = 1),
        legend.title=element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(x= "",
       y = "Number of target molecules",
       title = "Involment of upstream regulators in biological pathways (modules)",
       subtitle = "Before treatment ") +
  scale_fill_manual(values = c( "#F8766D",
                                "#ded031",
                                "#00BFC4")) 

modules_plot

ggsave(filename = "06_modules_plot.png",
       plot = last_plot(),
       device = "png",
       path = "../results",
       scale = 1,
       limitsize = TRUE)
               
                    
```

The plot indicates which regulator influences specific biological pathways. This information, together with the target molecules identification, implies that TGFB1, and possibly other molecules of the module 2 pathway, could be potential drug targets for RA treatment.
